---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(tidyr)
library(ggrepel)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```

* * *

## Part 1: Data

### Insights on data collection
- It appears the data was collected through an observational study.

- The method in which this data was collected appears to make it generalizable
to the non-institutionalized Adult population (over 18 years of age.)
- The conclusions arrived at cannot be determined as causal, as random assignment
  was not utilized.
  
- The observational study itself appears to follow a Stratified sampling method.
  - Each state can be considered a strata, and within that strata, random samples
  pertaining to the outlined requirements are completed. 
  - For the land-line telephone survey, random households are called within the 
  strata (state), and a randomly selected adult is interviewed. For the cellular 
  telephone version, data is collected from an adult who resides in a private 
  residence or college housing.
  
- In the data collection phase, there could be a few issues.
  - When collecting data from each state, a random sample of the state is taken,
  however, it's possible that the randomized sample happens to include data taken
  from households in specific regions of the state, that are either more or less
  wealthy. If too much data from those regions is acquired, it could easily skew 
  the outcome of the results. To ensure a more evenly distributed set of survey
  data is acquired, a more granular breakdown of random sampling would be needed,
  such as multi-stage sampling of each state (strata), so all regions, rich or poor,
  have an equal chance of being included in the survey.
  - There are a substantial more female interviees in the dataset than males.
  - Specific state quality of life metrics could skew the overall interview results
  if some are much lower than others. If attempting to identify trends among the
  sample, they could end up biased.
  - Up until 2001, the data collected up to that point would not be reliable, as
  many states were not taking part in the surveys.

* * *

## Part 2: Research questions

**Research question 1:**

## Which states are the healthiest on average?
- Variables: genhlth, X_state

- This surface-level question is aimed at identifying which states are the healthiest, based on the
genhlth and X_state variables.
- The goal is to provide a high-level indication of which states aer the healthiest, so further analysis
can be conducted as to why that may be the case.

**Research question 2:**

## Is there a relationship between overall health, income, and education?
- Variables: genhlth, income2, educa, sex.

- This question, working in conjunction with the first, is to try and answer if there is a relationship
between an individuals income, education and overall health. This would tie in with which states are on average,
the healthiest.

**Research question 3:**

## What is the average interview completion rate by state and communication
## device.
- Variables: dispcode, X_state, qstver.

- This question will answer which states had the highest interview completion 
rate.

* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

### Preliminary Data Analysis
```{r prelim_checks, message = FALSE}
head(brfss2013)
tail(brfss2013)
names(brfss2013)
str(brfss2013)
```
```{r cleaning, message = FALSE}
# Keep only relevant columns for analysis
trimmed_brfss2013 <- brfss2013 %>%
  select(c(X_state, genhlth, qstver, dispcode, sex, educa, income2))

colnames(trimmed_brfss2013)
View(trimmed_brfss2013)
glimpse(trimmed_brfss2013)

# Check N/A's
sapply(trimmed_brfss2013, function(x) sum(is.na(x)))

# Remove rows with N/A's
trimmed_brfss2013 <- na.omit(trimmed_brfss2013)

trimmed_brfss2013 <- rename(trimmed_brfss2013,
  state = X_state,
  gen_hlth = genhlth,
  int_type = qstver,
  compl_type = dispcode,
  sex = sex,
  education = educa,
  income = income2)

```

```{r prelim_analysis, message = FALSE}
# Check for biases
table(trimmed_brfss2013$sex)

# Compare sex and state
table(trimmed_brfss2013$sex,
trimmed_brfss2013$state)

# Compare interview type and completion state
table(trimmed_brfss2013$int_type,
trimmed_brfss2013$compl_type)

# Compare education level and income
table(trimmed_brfss2013$education,
trimmed_brfss2013$income)

# Compare income and state
table(trimmed_brfss2013$income,
trimmed_brfss2013$state)
```

**Research question 1:**

```{r}
# Comparing State and General Health
table(trimmed_brfss2013$state,
trimmed_brfss2013$gen_hlth)

trimmed_brfss2013 %>%
  group_by(gen_hlth) %>%
  summarise(count = n())

ggplot(data = trimmed_brfss2013, aes(y = state, color = gen_hlth)) +
  geom_bar() +
  theme(legend.position = "bottom") +
  labs(
    x = "Count", y = "State",
    title = "General Health by State",
    subtitle = "",
    caption = "",
    color='General Health'
  )

  ggsave(
  'General Health by State.png',
  plot = last_plot(),
  scale = 1,
  width = 8,
  height = 8,
  dpi = 500,
  limitsize = TRUE,
  bg = NULL
  )

  
```

**Research question 2:**

```{r}

```

**Research question 3:**

```{r}

```

