---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
View(brfss2013)
```



* * *

## Part 1: Data

### Insights on data collection
- It appears the data was collected through an observational study.

- The method in which this data was collected appears to make it generalizable
to the non-institutionalized Adult population (over 18 years of age.)

- The conclusions arrived at cannot be determined as causal, as random assignment
  was not utilized.
  
- The observational study itself appears to follow a Stratified sampling method.
  - Each state can be considered a strata, and within that strata, random samples
  pertaining to the outlined requirements are completed. 
  - For the land-line telephone survey, random households are called within the 
  strata (state), and a randomly selected adult is interviewed. For the cellular 
  telephone version, data is collected from an adult who resides in a private 
  residence or college housing.
  
- In the data collection phase, there could be a few issues.
  - When collecting data from each state, a random sample of the state is taken,
  however, it's possible that the randomized sample happens to include data taken
  from households in specific regions of the state, that are either more or less
  wealthy. If too much data from those regions is acquired, it could easily skew 
  the outcome of the results. To ensure a more evenly distributed set of survey
  data is acquired, a more granular breakdown of random sampling would be needed,
  such as multi-stage sampling of each state (strata), so all regions, rich or poor,
  have an equal chance of being included in the survey.

  - Specific state quality of life metrics could skew the overall interview results
  if some are much lower than others. If attempting to identify trends among the
  sample, they could end up biased.
  
  - Up until 2001, the data collected up to that point would not be reliable, as
  many states were not taking part in the surveys.


* * *

## Part 2: Research questions

**Research question 1:**

## Is income tied to overall health?

**Research question 2:**

## Which states are the healthiest? (Top 5?)

**Research question 3:**

## What are the largest factors to good health?


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research question 1:**

```{r}

```



**Research question 2:**

```{r}

```



**Research question 3:**

```{r}

```

