---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(tidyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```

* * *

## Part 1: Data

### Insights on data collection
#### Generalizability and Causality
- The method in which this data was collected (observational study) appears to make it generalizable
to the non-institutionalized Adult population (over 18 years of age.)
- The conclusions arrived at cannot be determined as causal, as random assignment
  was not utilized.

#### Sampling Method and Potential Biases
- The observational study itself appears to follow a Stratified sampling method.
  - Each state can be considered a strata, and within that strata, random samples are taken.
  
- In the data collection phase, there could be a few issues.
  - Bias could exist within the sampling method; more interviews may be taken in specific
  neighborhoods that are of higher or lower income, skewing the overall results. A multi-stage sampling
  method would likely return more consistent results.
  - There are a substantial more female interviees in the dataset than males.
  - Up until 2001, the data collected would not be reliable, as
  many states were not taking part in the surveys at that time. It took until 2011 before data
  from all states were being collected.

* * *

## Part 2: Research questions

### Research question 1

#### Which states are the healthiest on average?

- This surface-level question is aimed at identifying which states are the healthiest, based on the
genhlth and X_state variables. The sex variable is added to help identify if there is a gap that exists
between males and females.

### Research question 2

#### Is there a relationship between general health, income, education level and sex?

- This question, working in conjunction with the first, is to try and answer if there is a relationship
between an individuals income, education and overall health. This would tie in with which states are on average,
the healthiest.

### Research question 3

#### What is the completion rate of interviews by type and sex?

- This question aims to answer which interview type had the highest completion rate. 
The sex variable is also included to help differentiate between whether males or females
were more likely to finish the interview or not.

* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

### Preliminary Data Analysis
```{r prelim_checks, message = FALSE}
#head(brfss2013)
#tail(brfss2013)
#names(brfss2013)
#str(brfss2013)
```

```{r cleaning, message = FALSE}
# Keep only relevant columns for analysis
trimmed_brfss2013 <- brfss2013 %>%
  select(c(X_state, genhlth, qstver, dispcode, sex, educa, income2))

#colnames(trimmed_brfss2013)
#View(trimmed_brfss2013)
#glimpse(trimmed_brfss2013)

# Check N/A's
#sapply(trimmed_brfss2013, function(x) sum(is.na(x)))

# Remove rows with N/A's
trimmed_brfss2013 <- na.omit(trimmed_brfss2013)

trimmed_brfss2013 <- rename(trimmed_brfss2013,
  state = X_state,
  gen_hlth = genhlth,
  int_type = qstver,
  compl_type = dispcode,
  sex = sex,
  education = educa,
  income = income2)
```

```{r prelim_analysis, message = FALSE}
# Check for biases
table(trimmed_brfss2013$sex)

# Compare sex and state
table(trimmed_brfss2013$sex,
trimmed_brfss2013$state)

# Compare interview type and completion state
table(trimmed_brfss2013$int_type,
trimmed_brfss2013$compl_type)

# Compare education level and income
table(trimmed_brfss2013$education,
trimmed_brfss2013$income)

# Compare income and state
table(trimmed_brfss2013$income,
trimmed_brfss2013$state)

# Compare general health and state
table(trimmed_brfss2013$state,
trimmed_brfss2013$gen_hlth)
```

### Research question 1
```{r}
# Comparing State and General Health and Sex
#table(trimmed_brfss2013$state,
#trimmed_brfss2013$gen_hlth)

trimmed_brfss2013 %>%
  group_by(gen_hlth) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

gen_hlth_prop_all <- with(trimmed_brfss2013, table(trimmed_brfss2013$gen_hlth)) %>%
prop.table()
gen_hlth_prop_all

gen_hlth_prop_sep <- with(trimmed_brfss2013, table(state, gen_hlth)) %>%
prop.table(margin = 1)
gen_hlth_prop_sep

# Plot
ggplot(data = trimmed_brfss2013, aes(y = state)) +
  guides(col = guide_legend(reverse = TRUE)) +
  geom_bar(data = trimmed_brfss2013, aes(color = sex)) +
  facet_grid(. ~ gen_hlth) +
  theme(legend.position = "bottom") +
  labs(
    x = "Count", y = "State",
    title = "General Health by State",
    subtitle = "2013 BRFSS Data",
    caption = "Kyle Hollands - September 25th, 2022",
    color='Sex'
  )

  ggsave(
  'General Health by State.png',
  plot = last_plot(),
  scale = 1,
  width = 8,
  height = 8,
  dpi = 500,
  limitsize = TRUE,
  bg = NULL
  )
```

### Research question 2

```{r}
# Comparing General Health, Income, Education.
#table(trimmed_brfss2013$income,
#trimmed_brfss2013$gen_hlth,
#trimmed_brfss2013$education)

trimmed_brfss2013 %>%
  group_by(gen_hlth, income, education) %>%
  summarise(count = n()) %>%
  arrange(desc(income), desc(education), gen_hlth)

gen_hlth_vs_income_prop <- with(trimmed_brfss2013, table(income, gen_hlth)) %>%
prop.table(margin = 1)
gen_hlth_vs_income_prop

#gen_hlth_vs_education_prop <- with(trimmed_brfss2013, table(gen_hlth, education)) %>%
#prop.table(margin = 1)
#gen_hlth_vs_education_prop

# Plot
ggplot(data = trimmed_brfss2013, aes(y = income)) +
  guides(col = guide_legend(reverse = TRUE)) +
  geom_bar(data = trimmed_brfss2013, aes(color = education)) +
  facet_grid(gen_hlth ~ sex) +
  theme(legend.position = "bottom") +
  labs(
    x = "Count", y = "Income",
    title = "General Health vs Income, Education and Sex",
    subtitle = "2013 BRFSS Data",
    caption = "Kyle Hollands - September 25th, 2022",
    color=""
  )

  ggsave(
  'General Health vs Income and Education.png',
  plot = last_plot(),
  scale = 1,
  width = 11,
  height = 8,
  dpi = 500,
  limitsize = FALSE,
  bg = NULL
  )
```

### Research question 3

```{r}
# Comparing state, interview type and completion type

#table(trimmed_brfss2013$int_type,
#trimmed_brfss2013$compl_type)

#trimmed_brfss2013 %>%
#  group_by(compl_type, int_type) %>%
#  summarise(count = n())
  
compl_perc <- with(trimmed_brfss2013, table(trimmed_brfss2013$compl_type)) %>%
prop.table()
compl_perc

compl_perc_sex <- with(trimmed_brfss2013, table(sex, compl_type)) %>%
prop.table(margin = 2)
compl_perc_sex

#compl_perc_state <- with(trimmed_brfss2013, table( state, compl_type)) %>%
#prop.table(margin = 1)
#compl_perc_state

# Plot
ggplot(data = trimmed_brfss2013, aes(x = compl_type)) +
  guides(col = guide_legend(reverse = TRUE)) +
  geom_bar(data = trimmed_brfss2013, aes(color = int_type)) +
  facet_grid(. ~ sex) +
  theme(legend.position = "bottom") +
  labs(
    x = "Completion Type", y = "# of Interviees",
    title = "Interview Completion by type and Sex",
    subtitle = "2013 BRFSS Data",
    caption = "Kyle Hollands - September 25th, 2022",
    color="Interview Type"
  )

ggsave(
  'Interview Completion by type and Sex.png',
  plot = last_plot(),
  scale = 1,
  width = 8,
  height = 8,
  dpi = 500,
  limitsize = FALSE,
  bg = NULL
  )
```

